Type Checker Generation using Reference Attribute Grammars

### 2023-11-16 MSc.tex
Make various minor changes

\documentclass[nofilelist]{cslthse-msc}
% to show a list of used packages at the end of the document, delete the nofilelist option
%\documentclass{cslthse-msc}
\usepackage[utf8]{inputenc}
\usepackage[main=british,swedish]{babel}
\usepackage{amsmath}
%\usepackage{amsfonts}
%%\usepackage{amssymb}
\usepackage{amsthm}
%\usepackage{makeidx}
\usepackage{graphicx}
\usepackage[titletoc, header, page]{appendix}
\usepackage{transparent}
\usepackage{tikz}
\usepackage{enumitem}
\usepackage{listings}
\usetikzlibrary{fit,positioning,shapes.geometric}

@@

@@

\keywords{MSc, BSc, template, report, style, structure}

@@

\chapter{Background}
%\section{Theory} % should contain the theoretical background needed to understand your work, if necessary.
\section{Type systems}
\subsection{Data types}
\subsection{Type checking}
\subsubsection{Dynamic type checking}
\subsubsection{Static type checking}

\section{Reference Attribute Grammars}
\subsection{Reference Attribute Grammars}
\subsection{JastAdd}
\subsubsection{AST definitions}\label{astdef}

@@

\section{Architecture}
In this section we'll be providing an overview of our architecture.
Since it contains many different components, it may be helpful to refer to the overview graph in figure \ref{overviewgraph} while reading.
It represents the flow of the program, with components and artefacts (represented as rectangular boxes) with arrows linking them to actions (elliptical boxes) which in turn have arrows to the new components and artefacts they produce.
The dashed arrows from represent links from the template components that exist in the current experimental state of the project, but in a finished product would instead link to component from the object language's own compilation environment.
%It uses rectangular boxes to represent the concrete data or artefacts of the procedure, and elliptical boxes to represent the actions performed with them.
\begin{figure}[]
  \resizebox{\textwidth}{!}{%
  \begin{tikzpicture}[
      every node/.style = {draw=black, rounded corners, align=center},
      every path/.style = {draw, -latex}
    ]
    \node (typerules)                               {Typing rules\\definition (\ref{typingruledefinition})};
    \node (trparsing) [right=of typerules, ellipse] {Scan,\\parse,\\build};
    \node (trparser)  [above=of trparsing]          {Typing rule parser};
    \node (trast)     [right=of trparsing]          {Typing rules AST};

@@

    \node (outputtypechecking) [below=of typechecker, ellipse] {Perform\\typechecking};
    \node (inputast)           [left=of outputtypechecking]    {Input program\\AST};
    \node (outputparsing)      [left=of inputast, ellipse]     {Scan,\\parse,\\build};
    \node (typeerrors)         [right=of outputtypechecking]   {Type errors};
    \node (inputprog)          [left=of outputparsing]         {Input program};

@@

\subsubsection{Parser}
\subsubsection{Abstract syntax tree}

@@

\subsection{Output type checker}
\subsubsection{Tree-representation}
\subsubsection{Type checking algorithm}\label{typecheckingalgorithm}
Generate code for each node and typing rule
Simple algorithm
Recursive
\subsubsection{Error handling}

@@

\cite{Hedin2011}
\section{Implementation} % if your work involved building an artefact/implementation, give the details here. Note, that this should not, as a rule, be a chronological description of your efforts, but a view of the result. There is a place for insights and lamentation later on in the report, in the Discussion section.
%TODO
\subsection{Type checker generation}
\begin{figure}[h]
\begin{lstlisting}[]
RuleSet ::= Rule*;

@@

\begin{appendices} %should contain lengthy details of the experimental setup, mathematical proofs, code download information, and shorter code snippets. Avoid longer code listings. Source code should rather be made available for download on a website or on-line repository of your choosing.
Git repository link

>>>>

\documentclass[nofilelist]{cslthse-msc}
% to show a list of used packages at the end of the document, delete the nofilelist option
%\documentclass{cslthse-msc}
\usepackage[utf8]{inputenc}
\usepackage[main=british,swedish]{babel}
\usepackage{amsmath}
%\usepackage{amsfonts}
%%\usepackage{amssymb}
\usepackage{amsthm}
%\usepackage{makeidx}
\usepackage{graphicx}
\usepackage[titletoc, header, page]{appendix}
\usepackage{transparent}
\usepackage{tikz}
\usepackage{enumitem}
\usepackage{listings}
\usepackage{subcaption}
\usetikzlibrary{fit,positioning,shapes.geometric}

@@

\setcounter{secnumdepth}{4} % Number subsubsections

@@

%\keywords{MSc, BSc, template, report, style, structure}
\keywords{type checking, type theory, code generation}

@@

\chapter{Background}
%\section{Theory} % should contain the theoretical background needed to understand your work, if necessary.
\section{Type systems}
\subsection{Data types}
\subsection{Type checking}
\subsubsection{Dynamic type checking}
\subsubsection{Static type checking}
\subsection{Typing rules}
use left hand side of these again \ref{trexampleif}

\section{Reference Attribute Grammars}
\cite{RAG}
\subsection{Reference Attribute Grammars}
\subsection{JastAdd}
\subsubsection{AST definitions}\label{astdef}

@@

\section{Architecture}
In this section we'll be providing an overview of our architecture.
Since it contains many different components, it may be helpful to refer to the overview graph in figure \ref{overviewgraph} while reading.
It represents the flow of the program, with components and artefacts (represented as rectangular boxes) with arrows linking them to actions (elliptical boxes) which in turn have arrows to the new components and artefacts they produce.
The dashed arrows from represent links from the template components that exist in the current experimental state of the project, but in a finished product would instead link to component from the object language's own compilation environment.
%It uses rectangular boxes to represent the concrete data or artefacts of the procedure, and elliptical boxes to represent the actions performed with them.
\begin{figure}[]
  \resizebox{\textwidth}{!}{%
  \begin{tikzpicture}[
      every node/.style = {draw=black, rounded corners, align=center},
      every path/.style = {draw, -latex}
    ]
    \node (typerules)                               {Typing rules\\definition (\ref{typingruledefinition})};
    \node (trparsing) [right=of typerules, ellipse] {Scan,\\parse,\\build};
    \node (trparser)  [above=of trparsing]          {Typing rule parser\\(\ref{typingrulesparser})};
    \node (trast)     [right=of trparsing]          {Typing rules AST\\(\ref{typingrulesast})};

@@

    \node (outputtypechecking) [below=of typechecker, ellipse] {Perform\\typechecking};
    \node (inputast)           [left=of outputtypechecking]    {Input program\\AST};
    \node (outputparsing)      [left=of inputast, ellipse]     {Scan,\\parse,\\build};
    \node (typeerrors)         [right=of outputtypechecking]   {Type errors\\(\ref{errorhandling})};
    \node (inputprog)          [left=of outputparsing]         {Input program};

@@

\subsubsection{Parser}\label{typingrulesparser}
\subsubsection{Abstract syntax tree}

@@

\subsection{Output type checker}
\subsubsection{Tree-representation}
\subsubsection{Type checking algorithm}\label{typecheckingalgorithm}
Generate code for each node and typing rule
Simple algorithm
Recursive

@@

\section{Implementation} % if your work involved building an artefact/implementation, give the details here. Note, that this should not, as a rule, be a chronological description of your efforts, but a view of the result. There is a place for insights and lamentation later on in the report, in the Discussion section.

\subsection{Type checker generation}
\begin{figure}[h]
\begin{lstlisting}[]
RuleSet ::= Rule*;

@@

\begin{appendices} %should contain lengthy details of the experimental setup, mathematical proofs, code download information, and shorter code snippets. Avoid longer code listings. Source code should rather be made available for download on a website or on-line repository of your choosing.
\chapter{Source code}
\url{https://github.com/nichobi/thesis-project}
